---
title: "Week 1"
author: "Qizhe Cui"
date: "26/07/2020"
output:
  html_document: default
  pdf_document: default
---
1. A vector contains objects from the same class, it can be expressed as c() or vector("type", length = )
2. empty vector can be created by 
```{r}
vector("character", length = 3)
vector("logical", length = 4)
x = c(3,4,5,6,1)
as.logical(x)
```
## How to create a matrix

We have three methods listed below
```{r}
#Assign a matrix with value column_wise
m <- matrix(1:9, nrow = 3, ncol = 3)
print (m)

#Give a vector its dimension
n = 1:6
dim(n)= c(2,3)
print(n)

#binding several vector in cbinding or rbinding

a= 1:3
b=2:4
c = 3:5
d=4:6
rbind(a,b,c,d)
cbind(a,b,c,d)
```
# Factor

We can create a factor which represents categorical data and it is self-describing
```{r}
 p = factor(c("yes","yes","no","no", "yes", TRUE, 1 +3i, "Hello"), levels = c("yes", "no", "TRUE", 1+3i, "Hello"))
q = factor(c("yes","yes","no","no", "yes", TRUE, 1 +3i, "Hello"))
p
q
table(p)
unclass(p)
unclass(q)


```
 **So we sometimes need to redefine the levels of a factor to cater a baseline level**  
 
 
# Missing values

NaN: missing numeric value  

NA: missing value including integer NA, character NA etc  


NaN is a special case of NA
```{r}
x = c(NA, NaN)
is.na(x)
is.nan(x)

y= Inf/ Inf
y


```
# Data Frames

We can create a data frame or convert a data frame into a matrix. Unlike matrix, data frame can store different classes.  
We normally read a dataframe by **read.csv( ) or read.table( )**
```{r}
 x = data.frame("weather" = c("hot", "cold", "humid", "warm"), "bar" = c(TRUE, TRUE, TRUE, FALSE))

x
row.names(x)
data.matrix(x)

```

 # Names
 
 R objects can have name, for example a vector
```{r}
y<-1:4
names(y)
names(y) <- c("sushi","nigiri","ramen","katsu")
y
```
 
 For list to have names:
```{r}
z <- list(1,2,3,9)
P <- list(a=1, b =2, sushi =3, haha = 9)
z
P
```
For matrix to have name on its column and row:
```{r}
m = matrix(1:4, nrow = 2, ncol = 2)
dimnames(m)<-list(c("quality", "price"), c("和牛", "sushi"))
m
```
# deput and dump

If we have build a function or object and we want to use it agian later, then we can use the dput/dump function to save it to a file. Dump can be used for multiple objects
```{r}
Q= matrix(1:9, nrow = 3, ncol = 3)
W = list(a="sushi", b ="nigiri", c<-"ramen")
Q
dput(Q)
#dput(Q, file = "save Q for use.R")
#dump(c("Q", "R"), file ="both Q and R.R"))
#We cannot show it but this can save both Q and R
```
 **dget("") can only be used on a single object**  

# How to read data?

We can read data from csv file or directly from url and we can read only part of the data or compressed file.
```{r}
a = file("Fish .csv")
b =readLines(a,10)
c =read.csv("Fish .csv") #a is the same as c
d =url("https://www.jhsph.edu/")


```
we can also use **gzfile** or **bzfile** to open compressed file in each compressed form

# Subetting 

we can use [] or [[]] or $ to subset R objects
```{r}
x = c("a","a","b","ahaha",TRUE,FALSE,5,9, 3+4i)
x[1:4]
u <- x>"a" #u compare all elements in x withx "a"
u
x[u] #we get all elements in x that are greater than "a
```
# Subsetting a list 
We can subsetting a list that has names for its element. Note that the [[ ]] is cannot extract multiple elements with only one exception [[c() ]]. [[ ]] will return a 
```{r}
x = list(a= 3,b=4,c=5,d=6,fizz = 0)
x[1:3]
x[[1]]
x$fizz

y = list(one = list(1,3,5,6,2), two = list(3.14,2.98,8.63))
y[[1]][[2]]
y[[c(2,3)]] #The only expcetion that double braket can take multiple elements
y[[2]][1:2] #The first two elements of the second subset
jesus ="one"
y[[jesus]] #We can replace the name of an lelement for convinence


```
# How to subset a matrix

We can use similar method of list to subset a specific element or lines of a matrix
```{r}
x=matrix(1:9, 3,3)
x[3,]
x[,2]
x[,1, drop =FALSE] #drop = FALSE can return a matrix rather than a sequence of number
x[2,3, drop =FALSE]
```
## Partial matching 
```{r}
b = list(hahaha=TRUE, walada=3)
b$h #This can save a lot of typing 
b[["w", exact =FALSE]] #We need to cancel the exact typing to allow 
```
# Subsetting Removing missing values

Remove NA value in a list or in multiple lists
```{r}
a = list(first= NA, second = NA, third =TRUE, fourth = 4, fifth = 1 +3i)
bad <- is.na(a)
bad
a[!bad] #This preserve the part of list that does not have NA
```
 In order to remove NA in a database we need to use methods **good = complete.cases(a)** in which good is a variable denoting the complete cases in object a. Then we can use ** a[good] ** to obtain the data that don't have NA.


# vectorised operation
One advantage of R is that we can vectorised mathematical operation so we get can several result in one go. This includes sequence and matrix:
```{r}
x = c(1.4,5.3,2.5,9.1)
y = 2:5
x+y
x * y

```
```{r}
a = matrix (rep(10,9),3,3)
b=matrix(c(1,3,2,4,2,1,4,21,3), 3,3)
a
b
a + b
a *b
a %*% b #This is the real matrix multiplication

```

# Dataframe useful function

```{r}
#If we want to get the mean of each column and rule out the NA value
l = read.csv("hw1_data.csv")
colMeans(l,na.rm = TRUE)

#If we want to read certain rows

l[c(1,3,5),]

#If we want the mean of a single column

mean(l$Ozone, na.rm = TRUE) #We use $ to turn the dataframe into a sequence

mean(l[["Wind"]],na.rm = TRUE) #Note that [[ ]] return a sequence of number which is accepted by mean function

sort(l$Ozone, decreasing = TRUE) #Sort can only apply to one column each time

max(l$Ozone, na.rm = TRUE) #This return the max of a column

#If we want to select conditional rows then we use it which return the rows where month =5 

l[l["Month"] ==5,]



```



